---
title: "ch12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(forcats)
```

## General social survey

### 1. Explore the distribution of `rincome`.

```{r 1}
gss_cat %>% 
  count(rincome)

gss_cat %>% 
  ggplot(aes(rincome)) +
  geom_bar()
```

It is hard to understand because the order of the factors are messed.


### 2.  What is the most common `relig` in this survery? What is the most common `partyid`?

```{r 2}
gss_cat %>%
  count(relig) %>%
  arrange(desc(n))

gss_cat %>%
  count(partyid) %>%
  arrange(desc(n))

```

### 3. Which `relig` does `denom` apply to? How can you find out with a table? How can you find out with visualization?

```{r 3}
gss_cat %>%
  filter(denom != "Not applicable") %>%
  count(relig) %>%
  arrange(desc(n))

gss_cat %>%
  mutate(is_denom = (denom == "Not applicable")) %>%
  ggplot(aes(relig, fill = is_denom)) +
  geom_bar()

```

## Modifying factor order

### 1 Some TV hours are suspiciously high. Is the mean a good summary?
```{r 4}
gss_cat %>%
  filter(!is.na(tvhours)) %>%
  ggplot(aes(reorder(relig, tvhours), tvhours)) +
  geom_boxplot() +
  coord_flip() +
  labs(y = "relig")

```

Maybe a box-plot is better

### 2. Check the orders of the factors in `gss_cat`.

```{r 5}
fct_ind <- sapply(gss_cat, is.factor)
cats <- gss_cat[,fct_ind]
lapply(cats, levels)

```
Seems like it is arranged.

### 3. Why did moving "Not applicable" to the front of the levels?

It is the smallest value, when represented in integer. To reverse the order, 'fct_rev' can help.

```{r 6}
gss_cat$rincome %>% levels()

gss_cat$rincome %>% relevel("Not applicable") %>% levels()

gss_cat %>%
  count(relig, sort = T)

```

## Modifying factor level

### 1. Have the proportions of people identifying as Democrat, Republican, and Independent changed over time?

```{r 7}
gss_cat$partyid %>% unique()

gss_party <- gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  )) 

gss_party %>%
  filter(!is.na(age)) %>%
  group_by(age, partyid) %>%
  count() %>%
  group_by(age) %>%
  mutate(sum_n = sum(n),
    prop = n / sum_n*100) %>%
  ggplot(aes(age, prop, color = fct_reorder2(partyid, age, prop))) +
  geom_line() +
  labs(colour = "Party")

```

### 2. How do you collapse `rincome` to a small set of categories?

Bind the categories by the value ranges.












