---
title: "ch13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Creating Data/time

### 1. What happens if you parse a string that contains invalid data?

```{r 1}
ymd(c("2010-10-10", "bananas"))
```

It will show error in parse, and returns "NA".


### 2. What does the tzone argument to today() do? Why is it important?

```{r 2}
today(tzone = "GMT")
```

It is the timezone that are interested to find.

### 3. Use lubridate function to parse each of the following datas:

```{r 3}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14"

mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)
```

## Data-time components

### 1. How does the distribution of flight times within a day change over the course of the year?
```{r 4}
library(nycflights13)

flights %>%
  mutate(date_time = make_datetime(year, month, day, sched_dep_time  %/% 100, sched_dep_time %% 100)) %>%
  select(date_time, flight) %>%
  mutate(dep_hour = update(date_time, yday = 1),
         month = month(date_time, label = T)) %>%
  ggplot(aes(dep_hour, ..density.., color = month)) +
  geom_freqpoly()
  
```

The trend is almost the same!

### 2. Compare dep_time, sched_dep_time, and dep_delay. Are the consisistent? Explain your finding.
```{r 5}
convert_to_datetime <- function(year, month, day, time){
  make_datetime(year, month, day, time %/% 100, time %% 100)
}

flights %>%
  filter(!is.na(dep_time), !is.na(arr_time)) %>%
  mutate(sched_dep_datetime = convert_to_datetime(year, month, day, sched_dep_time),
         dep_datetime = convert_to_datetime(year, month, day, dep_time)) %>%
  select(sched_dep_datetime, dep_datetime, dep_delay, flight) %>%
  mutate(calcuated_delay = dep_datetime - sched_dep_datetime,
         difference = as.numeric(calcuated_delay/60 - dep_delay)) %>%
  ggplot(aes(difference)) +
  geom_histogram()
```

### 3. Compare air_time with the duration between the departure and arrival. Explain your findings.

```{r 6}
flights %>%
  filter(!is.na(dep_time), !is.na(arr_time), !is.na(air_time)) %>%
  mutate(
  dep_datetime = convert_to_datetime(year, month, day, dep_time),
  arr_datetime = convert_to_datetime(year, month, day, arr_time)
  ) %>%
  select(dep_datetime, arr_datetime, air_time, flight) %>%
  mutate(air_time_calculated = arr_datetime - dep_datetime,
  difference = as.numeric(air_time_calculated/60 - air_time)) %>%
  ggplot(aes(difference)) +
  geom_histogram(bins = 30)
```

### 4. How does the average delay time change over the course of a day?

```{r 7}
flights %>%
  filter(!is.na(dep_time), !is.na(arr_time), !is.na(air_time)) %>%
  mutate(sched_dep_datetime = convert_to_datetime(year, month, day, sched_dep_time)) %>%
  select(sched_dep_datetime, dep_delay, flight) %>%
  mutate(hour = hour(sched_dep_datetime)) %>%
  group_by(hour) %>%
  summarise(mean_delay = mean(dep_delay)) %>%
  ggplot(aes(hour, mean_delay)) +
  geom_line()
```

### 5. On what day of the week should you leave if you want to minimize the chance of delay?

```{r 8}
flights %>%
  filter(!is.na(dep_time), !is.na(arr_time), !is.na(air_time)) %>%
  mutate(sched_dep_datetime = convert_to_datetime(year, month, day, sched_dep_time)) %>%
  select(sched_dep_datetime, dep_delay, flight) %>%
  mutate(wday = wday(sched_dep_datetime, label = T)) %>%
  group_by(wday) %>%
  summarise(mean_delay = mean(dep_delay)) %>%
  ggplot(aes(wday,mean_delay)) +
  geom_bar(stat = "identity")
```

Saturday looks good!

### 6. What makes distribution of diamonds\$carat and flights\$sched_dep_time similar?

```{r 9}
ggplot(diamonds, aes(carat)) +
  geom_histogram(binwidth = 0.01)

ggplot(flights, aes(sched_dep_time)) +
  geom_histogram(binwidth = 3)

```
Rounding to whole numbers

### 7. Confirm the hypothesis that the early departures of flights in minutes 20-30 and 50-60 are caused by scheduled flights that leave early.

```{r 10}
breaks <- c(0:6)*10

flights %>%
  filter(!is.na(dep_time), !is.na(arr_time), !is.na(air_time)) %>%
  mutate(minute = sched_dep_time %% 100,
         minute = replace(minute, minute == 0, 60)) %>%
  select(minute, dep_delay) %>%
  mutate(is_delay = (dep_delay > 0),
         time_interval = cut(minute, breaks = breaks)) %>%
  group_by(time_interval) %>%
  summarise(delay_perc = sum(is_delay)/n())
  
```

## Time spans 

### 1. Why there is `months()` but no `dmonths()`?

Because the duration of each month is different.

### 2. Explain `days(overnight * 1)`, how does it work?

`T` and `F` will coerce to numerical.

### 3. Create a vector of dates giving the first day of every month in 2015. Create a vector giving the first day of every month in this year.

```{r 11}
ymd("20150101") + (0:11)*months(1)

ymd(paste0(year(today()),"01","01")) + (0:11)*months(1)
```

### 4. Write a program gives the years according to your time entry.

```{r 12}
find_year_old <- function(birthday){
  (ymd(birthday) %--% today()) %/% dyears(1)
}

find_year_old("20010201")
```

### 5. Why can't `today() %--% (today() + years(1)) / months(1)` work?

```{r 13}
(today() %--% (today() + years(1)) ) / months(1)
```
It works!



