---
title: "ch13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Creating Data/time

### 1. What happens if you parse a string that contains invalid data?

```{r 1}
ymd(c("2010-10-10", "bananas"))
```

It will show error in parse, and returns "NA".


### 2. What does the tzone argument to today() do? Why is it important?

```{r 2}
today(tzone = "GMT")
```

It is the timezone that are interested to find.

### 3. Use lubridate function to parse each of the following datas:

```{r 3}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14"

mdy(d1)
ymd(d2)
dmy(d3)
mdy(d4)
mdy(d5)
```

## Data-time components

### 1. How does the distribution of flight times within a day change over the course of the year?
```{r 4}
library(nycflights13)

flights %>%
  mutate(date_time = make_datetime(year, month, day, sched_dep_time  %/% 100, sched_dep_time %% 100)) %>%
  select(date_time, flight) %>%
  mutate(dep_hour = update(date_time, yday = 1),
         month = month(date_time, label = T)) %>%
  ggplot(aes(dep_hour, ..density.., color = month)) +
  geom_freqpoly()
  
```












